---
title: "Lab_3"
author: "Natarajan Shankar"
date: "July 27, 2016"
output: pdf_document
---

```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
# Load required libraries
library(gmodels)
library(ggplot2)
library(car)
library(lsr)
```

```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
# Load the base 1993 data provided along with assignment handout
load("GSS.Rdata")

```
\pagebreak

# \textcolor{blue}{DATA ANALYSIS AND SHORT ANSWER SECTION}

## \textcolor{blue}{=============================================================================}
## \textcolor{blue}{Problem 14: Conduct a Chi-Square test to determine if there is association}
## \textcolor{blue}{between marital status (marital) and political orientation (politics)}
## \textcolor{blue}=============================================================================}
```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
# extract just the two fields of interest - marital and politics - so that we can 
# deal with a smaller data frame. Call the data frame - dataDF
dataDF <- data.frame(GSS$marital, GSS$politics)
names(dataDF) <- c("marital", "politics")

# =============================================================================
# View the data to understand its makeup and quirks
# =============================================================================
# Check class type of "marital", verify as "factor" for applying chi square test
# Check class type of "politics", verify as "factor" for applying chi square test
class(dataDF$marital); class(dataDF$politics)

# Check levels of "marital" and "politics"
levels(dataDF$marital) ; levels(dataDF$politics)

# Also look at summary to see NAs
summary(dataDF$marital); summary(dataDF$politics)

# =============================================================================
# Clean the data
# =============================================================================
# Per the codebook, marital answers of "no answer" are included. Remove them
dataDF <- dataDF[!dataDF$marital == "no answer",]

# "marital" attribute has NA as factor, clean it up prior to processing
# "politics" has <NA>, these are true NOT available, remove them.
#dataDF <- dataDF[dataDF$marital != "NA",]
dataDF <- dataDF[-which(dataDF$marital == "NA"), ]
dataDF <- dataDF[!is.na(dataDF$politics),]

# Drop the "NA"factor from marital, those rows have been removed
dataDF$marital = droplevels(dataDF$marital)

# Check how many data points are in the base data frame
nrow(dataDF)

# =============================================================================
# Run the Chisquare test, also display CrossTable to verify expected cell count
# =============================================================================
# Run the Chisquare test
CrossTable(dataDF$marital, dataDF$politics, chisq=TRUE, expected=TRUE, sresid=TRUE, format="SPSS")
```
  * \textcolor{blue}{All cells show expected value above 5}
  * \textcolor{blue}{There is not a need to run the Fisher test, Chi square test is appropriate}

```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
# Conduct an effect size calculation
effect_size <- sqrt(44.2255/(nrow(dataDF) * 4))
effect_size
```
## =============================================================================
## Answers to questions : Problem 14
## =============================================================================
## Answer to Question 14 - A
  * Null Hypothesis : "Marital" and "Politics" are independent
  * Alternative Hypothesis : Gain confidence that "Marital" and "Politics" are related in some way

## Answer to Question 14 - B
    * Value of the test statistic is : 44.2255
    * p value is : 0.0001822704

## Answer to Question 14 - C
    * The effect size is : ```r  effect_size ```

  * Verify using lsr package also: ```r cramer_data <- cramersV(dataDF$marital, dataDF$politics); cramer_data```
  
## Answer to Question 14 - D
  * Conclusion:
  * There was an significant association between "marital" and "politics" with the test statistic of 44.225 at p < 0.05
  * This seems to represent that politics is impacted by marital 
  * Based upon this data, the NULL hypopthesis is rejected and confidence is gained that the alternative hypothesis is true.
  * Based upon the computed effect size of 0.08756363, the effect size is "small", leaning to "medium", for 16 degrees of freedom

\pagebreak

## \textcolor{blue}{=============================================================================}
## \textcolor{blue}{Problem 15: Conduct a Pearson Correlation Analysis to examine the association }
## \textcolor{blue}{between age when married (agewed) and hours of TV watched (tvhours)}
## \textcolor{blue}=============================================================================}
```{r, error=TRUE, fig.width = 3.5, fig.height = 3.5}

# extract just the two fields of interest
dataDF <- data.frame(GSS$agewed, GSS$tvhours)
names(dataDF) <- c("agewed", "tvhours")

# =============================================================================
# View the data to understand its makeup and quirks
# =============================================================================
# Check class type of "agewed" 
# Check class type of "tvhours"
class(dataDF$agewed); class(dataDF$tvhours)

# Both variables verified to be numeric

# Check data types in "agewed" and "tvhours"
summary(dataDF$agewed); summary(dataDF$tvhours)

# =============================================================================
# Clean the data
# =============================================================================
# The "no answer" for agewed is coded as 99, remove these
# the "no answer" for tvhours is also coded as 99, remove these
dataDF <- dataDF[!dataDF$agewed == 99,]
dataDF <- dataDF[!dataDF$tvhours == 99,]

# agewed value of 0 appears to be erroneous, remove them
dataDF <- dataDF[!dataDF$agewed == 0,]

# Re-check data types in "agewed" and "tvhours"
summary(dataDF$agewed); summary(dataDF$tvhours)

# Check how many data points are in the base data frame
nrow(dataDF)
# =============================================================================
# Get a feel for Normality of data
# =============================================================================
qqnorm(dataDF$agewed)
qqnorm(dataDF$tvhours)

# =============================================================================
# Run the Pearson test, assume that sample size is large enough to not
# require the running od Spearman test
# =============================================================================
pearson_result <- cor.test(dataDF$agewed, dataDF$tvhours, method="pearson", conf.level=0.95, na.rm=TRUE)
pearson_result

```

## =============================================================================
## Answers to questions : Problem 15
## =============================================================================
## Answer to Question 15 - A
  * Null hypothesis is  that correlation is 0, i.e no correlation between agewed and tvhours
  * Alternative hypothesis is that the correlaton between agewed and tvhours is different from 0
  
## Answer to Question 15 - B
  * Test statistic is : ```r round(pearson_result$statistic, 2)``` 
  * p-value is : ```r pearson_result$p.value```
  * Correlation coefficient is ```r sub("^(-?)0.", "\\1.", sprintf("%.2f", pearson_result$estimate))```

## Answer to Question 15 - C
* Conclusion:
  * Pearson's correlation between agewed and tv hours is ```r pearson_result$estimate``` and p-value is ```r pearson_result$p.value```
  * Based upon the correlation statistic and the statistically non-significant p value it can be conlcuded that agewed and tvhours are weakly correlated and that the Null hypotheses of zero correlation cannot be rejected

\pagebreak



## \textcolor{blue}{=============================================================================}
## \textcolor{blue}{Problem 16: Create a new binary variable "Married" that denotes whether an }
## \textcolor{blue}{individual is currently married or not married. Then consider just the 23 year }
## \textcolor{blue}{olds in this sample. Conduct a Wilcox ran-sum test to determine whether the }
## \textcolor{blue}{new married variable is associated with children of 23 year olds}
## \textcolor{blue}=============================================================================}

```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
# Look only for 23 year olds
dataDF <- GSS[GSS$age == 23,]

# For convenience, extract just the two fields of interest
dataDF <- data.frame(dataDF$marital, dataDF$childs)
names(dataDF) <- c("marital", "childs")

# =============================================================================
# View the data to understand its makeup and quirks
# =============================================================================
# Check class type of "agewed" - should be numeric
# Check class type of "tvhours" - should be factor
class(dataDF$marital); class(dataDF$childs)

# Check data types in "marital" and "childs"
summary(dataDF$marital); summary(dataDF$childs)

# =============================================================================
# Clean the data
# =============================================================================
# "marital" attribute has NA, clean it up prior to processing
dataDF <- dataDF[dataDF$marital != "NA",]

# Drop the "NA"factor from marital, those rows have been removed
dataDF$marital = droplevels(dataDF$marital)

## Create a new variable "married"
dataDF$married <- factor(ifelse(dataDF$marital == "married", as.numeric(1), as.numeric(0)))

# From the codebook, childs = 9 is a "no answer". Remove these rows
dataDF <- dataDF[!dataDF$childs == 99,]

# For convenience, extract just the two fields of interest
dataDF <- data.frame(dataDF$married, dataDF$childs)
names(dataDF) <- c("married", "childs")

# Check data types in "marital" and "childs"
class(dataDF$married); class(dataDF$childs)

# Check homogeniety of variance
leveneTest(dataDF$childs, dataDF$married, center=mean)
# p-value is statistically significant, The Homogeneity of variance requirement is violated
# However, the sample size is large, let's continue with Wilcox Rank Sum test

# =============================================================================
# Run the Wilcox Rank Sum test, ignore the warning message
# =============================================================================
# run the wilcox test
wilcoxModel <- wilcox.test(dataDF$childs, as.numeric(dataDF$married), paired=FALSE)
wilcoxModel

# What is the mean of your new "married" variable, proportion of cases coded 1, out of total
mean_married <- sum(dataDF[dataDF$married == "1",]$childs)/length(dataDF[dataDF$married == "1",]$childs)

# Proportion of married 23 year olds with children, amongst all with children
m_23_children <- nrow(dataDF[((dataDF$married == 1) & (dataDF$childs != 0)), ])/
                        nrow(dataDF[dataDF$married == 1, ])
```
## =============================================================================
## Answers to questions : Problem 16
## =============================================================================
## Answer to Question 16 - A
  * Mean of the new "married" variable is: ```r mean_married``` children per married 23 year old
  * Proportion of married 23 year olds with children : ```r m_23_children```
  
## Answer to Question 16 - B
  * Null hypothesis : The two groups (married and childs) are not different
  * Alternative hypothesis : The two groups (married and childs) are different

## Answer to Question 16 - C
  * Test statistic is ```r wilcoxModel$statistic```
  * p-value is ```r wilcoxModel$p.value```
  
## Answer to Question 16 - D : effect size calculation
```{r, error=TRUE, fig.width = 6, fig.height = 4.5}
rFromWilcox <- function (wilcoxModel, N) {
  z <- qnorm(wilcoxModel$p.value/2)
  r <- z/sqrt(N)
  cat(wilcoxModel$data.name, "Effect size, r = ", r)
}

rFromWilcox(wilcoxModel, nrow(dataDF))

```

## Answer to Question 16 - E  : Conclusion
  * Conclusion: Based upon the p-value and the statistic above, the two groups are different
  * and the NULL hypothesis can be rejected and confidence in alternate is gained
  * Married status is associated with number of children for respondents who are 23 years old. 
  * The computed effect size is subjectively "huge".

\pagebreak



## \textcolor{blue}{=========================================}
## \textcolor{blue}{Problem 17: Conduct an ANOVA to determine if there is an association between  }
## \textcolor{blue}{religious affiliation (relig) and age when married (agewed)}
## \textcolor{blue}{=========================================}
```{r, error=TRUE, fig.width = 6, fig.height = 4.5}

# extract just the fields of interest, "relig", "age", and "agewed"
# put into a convenience dataframe, dataDF
dataDF <- data.frame(GSS$relig, GSS$age, GSS$agewed)
names(dataDF) <- c("relig", "age", "agewed")

# =============================================================================
# View the important data to understand its makeup and quirks
# =============================================================================
# Check class type of "relig" - should be factor
# Check class type of "agewed" - should be factor
class(dataDF$relig); class(dataDF$agewed)

# Check data types in "relig" and "agewed"
summary(dataDF$relig); summary(dataDF$agewed)

# =============================================================================
# Clean the data
# =============================================================================
#Remove instances where agewed is greater than age
dataDF <- dataDF[!(GSS$agewed > GSS$age), ]

# remove the age column now, we are done with it
dataDF <- data.frame(dataDF$relig, dataDF$agewed)
names(dataDF) <- c("relig", "agewed")

# The "no answer" for agewed is coded as 99, remove these as well
dataDF <- dataDF[!is.na(dataDF$relig),]
dataDF <- dataDF[!dataDF$agewed == 99,]

# agewed value of 0 appears to be erroneous, remove them
dataDF <- dataDF[!dataDF$agewed == 0,]

# get rid of the NAs in the relig column
dataDF <- dataDF[-which(dataDF$relig == "NA"), ]

# get rid of the DKs in the relig column
dataDF <- dataDF[-which(dataDF$relig == "DK"), ]

# Drop the "NA"factor from relig, those rows have been removed
dataDF$relig = droplevels(dataDF$relig)

# =============================================================================
# Run Levene's test, Homogenity of variance is a requirement
# =============================================================================
leveneTest(dataDF$agewed, dataDF$relig, center = mean)
# The Levene Test p-value is not statistically significant, the homogenity of variance assumption 
# is not violated

# =============================================================================
# now run ANOVA and gather the output
# =============================================================================
aovData <- aov(agewed ~ relig, data = dataDF)
summary(aovData)

# Manually compute the effect size
effectSize <- (804 - (2* 24.75))/(804 + 29430 + 24.75)
effectSize ^ 2
```
## ==================================
## Answers to questions : Problem 17
## ==================================
## Answer to Question 17 - A
  * Null hypothesis : All group means are equal
  * Alternative hypothesis : At least one group mean is different from another
  
## Answer to Question 17 - B
  * The test statistic is an F value of : 8.12
  * The p value is : 1.85e-06
  
## Answer to Question 17 - C
  * There is statistically significant difference between individual pairs of groups
  * This is indicated by the p-value of < 0.05
  
## Answer to Question 17 - D
  * Conclusions:
  * The F ratio has the value 8.12, degrees of freedom (4, 1189), F(4, 1189) = 8.12
  * The F ratio value > 1 is indicating that all group means are not equal, This along
  * the p-value indicates that the Null hypothesis 
  * can be rejected and confidence in the alternative is gained. However, the effect 
  * size will determine actual practical significance
  * The effect  size is ```r effectSize^2```
  * In spite of statistically significanct results, the effect size is "low""

