---
title: "W203_Supplemental_4_I"
author: "Natarajan Shankar"
date: "July 20, 2016"
output: pdf_document
---

```{r, error=TRUE, fig.width = 8, fig.height = 4.5}
# Natarajan Shankar
# W 203: Supplement Exercie 4
# July 20th

# Problem 1: BOOTSTRAPPING
#

# Play it safe and set a seed value to get repeatable results
set.seed(314156)

# Pull in needed libraries
library(ggplot2)

# Chose number of samples to be 1001. An odd number helps read the meadian vector easie
NBS <- 1001

# Set the confidence interval (CI) at plus/minus 95% (or 0.05 total outside the CI), 
# split evenly across two tails i.e. 0.025 at each tail
LOW_CI <- 0.05/2
HIGH_CI <- 1 - LOW_CI

# FUNCTION: bootstrap_median()
# Inputs: Number of iterations, NBS and  the parent sample
# Iteration processing to extract many bootstrap samples 
bootstrap_median <- function (iteration, parent_sample) {
  
  # Take a child sample from the parent sample, with replacement.
  # Child sample has the same length as the parent sample
  child_sample <- sample(parent_sample, size = length(parent_sample), replace = TRUE)
    
  # Calculate the median of the new sample
  child_median <- median(child_sample)
  
  # return the calculated mean
  return(child_median)
}


# Main bootstrap processing function
#   1. Takes a parent sample (from one of Random Normal, Chi Square, Binomial)
#   2. Takes the number of bootstrap iteraions (NBS) 
#   3. Iterates over each sample and extracts a median
#   4. Returns a collection of bootstrapped medians and CI interval
median_bootstrap <- function(parent_sample, NBS) {
  
  # Compute the median across NBS sub samples
  median_distribution <- sapply(1:NBS, bootstrap_median, parent_sample = parent_sample)
  
  # Sort the medians from high to low
  median_distribution <- sort(median_distribution)
  
  # compute the 95% confidence interval
  low_CI <- median_distribution[round(NBS * LOW_CI)]
  high_CI <- median_distribution[round(NBS * HIGH_CI)]
  
  # Function returns the median distribution and confidence intervals
  return(list(median_distribution, c(low_CI, high_CI)))
}


# Test the median bootstrap using 3 kinds of data distributions
# 1. Random Normal
# 2. Poisson
# 3. Chi Square
# Plot results for each
test_code <- function() {
  # Random Normal Distribution
    parent_sample <- rnorm(n=100, mean=0, sd =1)
    results <- median_bootstrap(parent_sample, NBS)
    distribution_hist(results, "Random Normal Distribution")
    
  # Poisson Distribution
    parent_sample <- rpois(300, 4)
    results <- median_bootstrap(parent_sample, NBS)
    distribution_hist(results, "Poisson Distribution")
    
  # Chi Square Distribution
    parent_sample <- rchisq(300, df=7)
    results <- median_bootstrap(parent_sample, NBS)
    distribution_hist(results, "Chi Square Distribution")
}



# Use one common plot routine that can be used by the test code. 
# Customize the header strings so that each plat is distinguished by title
distribution_hist <- function (results, type = "") {
    resultsDF <- data.frame(results[[1]])
    myPlot <- ggplot(resultsDF, aes(results[[1]]))
    myPlot <- myPlot + geom_histogram(binwidth=0.02)
    myPlot <- myPlot + 
      ggtitle(paste ("Median Distribution (Samples from \n", type, ")", 
                "\nBlue lines show 95% confidence boundaries\n", "red is Median")) + 
                labs(x="Median Value", y="Frequency") +
                geom_vline(xintercept = results[[2]][1], color = "blue") + 
                geom_vline(xintercept= results[[2]][2], color="blue") +
                geom_vline(xintercept= results[[1]][round(1001/2)], color="red")
    plot(myPlot)
}

# Test the code above
test_code()
```

## Median distribution of samples with the Normal Distribution as parent show equal distribution around Median and are not clustered
## Median distribution of samples with the Poisson distribution as parent show binary values as expected, clustered around the Median
## Median distribution of samples with the Chi Square distribution as parent also shows even distribution around Median, not clustered


\pagebreak
```{r error=TRUE, fig.width = 8, fig.height = 4.5}
    
# Problem 2: OPTIMIZATION PROBLEM
#

# Function : isEmpty()
# Support routine to determine whether a vector is empty
# Used after cleaning up negative numbers fromm user input.
# If after cleanup of negative numbers, the vector is null, do not process further
isEmpty <- function(x) {
  return(length(x)==0)
}


# Function: objf()
# Scaffolding Cost function for Optim
# Take 3 numbers, x, r, and k. Raise r to the power of k and compare with X
# Return square of the compare

objf <- function(r, numbers, k) {
  
  cost = (numbers - (r^k)) ^2

  return(cost)
}

# Function: rootk()
# Function to find the root k of a positive number using numerical optimization
# Accept a vecor of numbers, numbers
# Return root k for all the numbers in the vector
rootk <- function(numbers, k, start = NULL) {
  
  if (!isEmpty(x <- numbers[numbers < 0])) {
    cat("\nCaution 0001 :  negative numbers in vector, being removed... \n\n\n")
    numbers <- numbers[numbers >= 0]

    # Catch illegal k values (does not work via rmarkdown, needs investigation!!)
    if (k < 0) {
      stop("\nERROR 001: Root request is invalid (below 0), Processing stopped")
    }
      
    # Stop if no numbers to process after negative number cleanup
    if(isEmpty(numbers))
        stop("\nERROR 002: Input had negative numbers, 
                    \ncleaned up number vector is empty!, Processing stopped")
  }
  
  # In case the user does not provide a starting approximation, start at 1
  is.null(start)
    start = 1
   
  # User vector has positive numbers and has been sanitized
  # apply Optim() to all members in the vector. Return a vecor of results
  # Use method BGFS for optimized processing
  # 
  roots <- sapply(numbers, function(i) {optim(par = i, 
                            numbers = i, objf, k = k, start = start, method="BFGS", control = list(maxit = 400))$par})
    result <- list(numbers, roots)
  
  # For pretty printing of final message
  if (k ==1)
      tag <- "st"
  else if (k ==2)
      tag <- "nd (square)"
  else if (k == 3)
      tag <- "rd (cube)"
  else 
      tag <- "th level"
  
    names(result) <- c("Processed Numbers", paste ("Corresponding", 
                          paste(k, tag, sep =""), "Root for each processed number"))
    return(result)

}
```

## \textcolor{blue}{Test case 1 : One element vector}
```{r error=TRUE}
rootk(4, 2)
```
## \textcolor{red}{- Note the negative root, accurate but interesting, larger start value does not change convergence on negative root}
## \textcolor{blue}{Test case 2 : two element vector}
```{r error=TRUE}
rootk(125,3, start = NULL)
```
## \textcolor{blue}{Test case 3 : One element, large number, big root}
```{r error=TRUE}
rootk(134217728,9)
```
## \textcolor{blue}{Test case 4 : 3 element vector, one negative}
```{r error=TRUE}
rootk(c(25, -125, 96), 3)
```
## \textcolor{blue}{Test case 5 : 3 element vector, negative root request}
### \textcolor{red}{"STOP" WORKS VIA CONSOLE, DOES NOT WORK VIA RMARKUP, NEEDS INVESTIGATION}
```{r error=TRUE}
rootk(c(25, 125, 96), -3)
```
## \textcolor{blue}{Test case 6 : 3 element vector, all 3 negative}
```{r error=TRUE}
rootk(c(-25, -125, -96), 5)

